Can Cultural Transmission Explain The Evolution of Cooperative Behavior?

### 2020-04-05 ms/main.tex
typo, reformatting, headings and paragraphs

@@

\textbf{Result 1: Vertical and horizontal transmission of cooperation.}
If 
\begin{equation}\label{eq:unequal_transmission}
c \cdot (1-T_B) < b \cdot  \alpha T_A  + (T_A - T_B) [1 + bp(1-\alpha)]
\end{equation}
then $\tilde{p}' > \tilde{p}$, and the frequency of phenotype $A$ among parents increases.

@@

%(yr: I stopped here)

\underline{Special case 2 - $\alpha =1$:} in this case, horizontal transmission can only occur as a result of interaction.
\\Eq.~\ref{eq:unequal_transmission} becomes:
\begin{equation} 
\begin{split} \label{eq:horizontal_transmission_only_after_interaction}
c\cdot(1-T_B)  < \,& T_A\cdot(1+b) - T_B
\end{split}
\end{equation}
Therefore, cooperative behavior will take over the population when eq.~\ref{eq:horizontal_transmission_only_after_interaction} is satisfied. 
This result is in accordance with the result presented by~\citet{lewin2017microbes}. 
\\\underline{Special case 3 - $\alpha = 0$:} in this case horizontal transmission can not occur as a result of interaction.  
Then, eq.~\ref{eq:unequal_transmission} becomes:
\begin{equation} 
\begin{split} \label{eq:horizontal_transmission_from_population}
c\cdot(1-T_B) < & (T_A-T_B)(1+bp)  
\end{split}
\end{equation}
From eq.~\ref{eq:horizontal_transmission_from_population} we learn that when $T_B\leq T_A$ cooperation will not take over the population.
When $T_A > T_B$ eq.~\ref{eq:horizontal_transmission_from_population} becomes:
\begin{equation} 
\begin{split} \label{eq:horizontal_transmission_from_population_reorder}
\frac{c\cdot(1-T_B)+T_B-T_A}{b\cdot(T_A-T_B)} < & p  
\end{split}
\end{equation}
Therefore, in the case $T_A>T_B$ a sufficient condition for cooperation fixation is:
\begin{equation} 
  \begin{split} \label{eq:sufficient_horizontal_transmission_from_population}
c < &\, \frac{T_A-T_B}{1-T_B}  
  \end{split}
  \end{equation}
\\So far, we assumed that $v=1$. The next section discuss the effect of oblique transmission ($v\neq 1$) which is common in cultural evolution models. 
Start by looking on special case where $v=0$. In this case eq.~\ref{eq:horizontal} becomes $\hat{p}=p$ and eq.~\ref{eq:nextgen_parents} becomes:
\begin{equation} 
\begin{split}\label{eq:nextgen_parents_oblique_only}
p' 
& = p^2 (1 - (1-p) (1-\alpha) T_B) \\
& + p(1-p)(p (1-\alpha) + \alpha) T_A \\
& + p(1-p)((1-\alpha)T_B + (1-\alpha)(p+(1-p)(1-T_B))) \\
& + (1-p)^2 p (1-\alpha) T_A
\end{split}
\end{equation}
\textbf{Result 2: Oblique and horizontal transmission of cooperation.} The frequency of phenotype A among adults increases if $p'>p$. From eq.~\ref{eq:nextgen_parents_oblique_only} applies that $p'>p$ is satisfied when:
\begin{equation} \label{eq:oblique_only_result}
T_A>T_B  
\end{equation}

>>>>

\subsection*{Without oblique transmission, $v=1$}

@@

\textbf{Result 1: Vertical and horizontal transmission of cooperation.}
If 
\begin{equation}\label{eq:unequal_transmission}
c \cdot (1-T_B) < b \cdot  \alpha T_A  + (T_A - T_B) [1 + bp(1-\alpha)]
\end{equation}
then $\tilde{p}' > \tilde{p}$, and the frequency of phenotype $A$ among parents increases. 
\\

@@

\paragraph*{Complete correlation between transmission and cooperation, $\alpha=1$.}

In this case, horizontal transmission can only occur as a result of cooperative interactions.
\\Eq.~\ref{eq:unequal_transmission} becomes:
\begin{equation} 
\begin{split} \label{eq:horizontal_transmission_only_after_interaction}
c\cdot(1-T_B)  < \,& T_A\cdot(1+b) - T_B
\end{split}
\end{equation}
Therefore, cooperative behavior will take over the population when eq.~\ref{eq:horizontal_transmission_only_after_interaction} is satisfied. 
This result is in accordance with the result presented by~\citet{lewin2017microbes}.

\paragraph*{No correlation between transmission and cooperation, $\alpha=0$.}

In this case horizontal transmission can not occur as a result of interaction.  
Then, eq.~\ref{eq:unequal_transmission} becomes:
\begin{equation} 
\begin{split} \label{eq:horizontal_transmission_from_population}
c\cdot(1-T_B) < & (T_A-T_B)(1+bp)  
\end{split}
\end{equation}

Therefore, since $c>0$, cooperation will not take over the population (and furthermore will become extinct) if $T_A\leq T_B$.
When $T_A > T_B$ eq.~\ref{eq:horizontal_transmission_from_population} becomes:
\begin{equation} 
\begin{split} \label{eq:horizontal_transmission_from_population_reorder}
p & > \frac{c\cdot(1-T_B)+T_B-T_A}{b\cdot(T_A-T_B)} .
\end{split}
\end{equation}
Hence, cooperation will take over the population when when eq.~\ref{eq:horizontal_transmission_from_population_reorder} is satisfied, that is if $p$ is above the threshold on the RHS. 
Specifically, a sufficient condition for cooperation fixation is found by setting the RHS to be negative:
\begin{equation} 
  \begin{split} \label{eq:sufficient_horizontal_transmission_from_population}
c < &\, \frac{T_A-T_B}{1-T_B}  
  \end{split}
  \end{equation}

\subsection*{With oblique transmission, $v<1$}

So far, we assumed that $v=1$. The next section discuss the effect of oblique transmission: transmission from non-parental adults.

\paragraph*{Without vertical transmission, $v = 0$.}

In this case eq.~\ref{eq:horizontal} becomes $\hat{p}=p$ and eq.~\ref{eq:nextgen_parents} becomes:
\begin{equation} 
\begin{split}\label{eq:nextgen_parents_oblique_only}
p' 
& = p^2 (1 - (1-p) (1-\alpha) T_B) \\
& + p(1-p)(p (1-\alpha) + \alpha) T_A \\
& + p(1-p)((1-\alpha)T_B + (1-\alpha)(p+(1-p)(1-T_B))) \\
& + (1-p)^2 p (1-\alpha) T_A
\end{split}
\end{equation}

\textbf{Result 2: Oblique and horizontal transmission of cooperation.} The frequency of phenotype $A$ among adults increases if $p'>p$.
From eq.~\ref{eq:nextgen_parents_oblique_only}, this occurs when:
\begin{equation} \label{eq:oblique_only_result}
T_A>T_B  
\end{equation}
